---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```


# NDTr: The Neural Decoding Toolbox in R


<!-- badges: start -->

[![Travis build status](https://travis-ci.com/emeyers/NDTr.svg?branch=master)](https://travis-ci.com/emeyers/NDTr)

[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/emeyers/NDTr?branch=master&svg=true)](https://ci.appveyor.com/project/emeyers/NDTr)

[![Coveralls test coverage](https://coveralls.io/repos/github/emeyers/NDTr/badge.svg)](https://coveralls.io/r/emeyers/NDTr?branch=master)

<!-- badges: end -->
  
<p>


## Overview

*Neural decoding* is a data analysis method that uses pattern classifiers to predict experimental conditions based on neural activity. The Neural Decoding Toolbox in R (NDTr) makes it easy to do neural decoding analyses in R. 


## Installation

You can install NDTr from github using:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("emeyers/NDTr")
```



## Usage

The package is based on 5 abstract object types: 

1. `Datasources (DS)`: generate training and test sets.
2. `Feature preprocessors (FP)`: apply preprocessing to the training and test sets.
3. `Classifiers (CL)`: learn relationships on the training set and make
predictions on the test data.
4. `Result Metrics (RM)`: summarize the prediction accuracies.
5. `Cross-validators (CV)`: take the DS, FP and CL objects and run a
  cross-validation decoding procedure.

By combing different versions of these 5 object types together, it is possible
to run a range of different decoding analyses.

Below is a brief illustration of how to use the NDTr to do a simple decoding
analysis. To learn how to use the NDTr please see the [introduction
tutorial](introduction_tutorial.html), the [generalization
tutorial](generalization_tutorial.html) and the [documentation
website](https://emeyers.github.io/NDTr/).



```{r basic_ndtr_analysis}
library(NDTr)

# file to data in "binned format"
basedir_file_name <- system.file(file.path("extdata", "ZD_150bins_50sampled.Rda"), package="NDTr")

# create the DS, FP, CL, RM, and CV objects
ds <- ds_basic(basedir_file_name, 'stimulus_ID', 5, num_label_repeats_per_cv_split = 3)
fps <- list(fp_zscore())
cl <- cl_max_correlation()
rms <- list(rm_main_results(aggregate_normalized_rank = "diag"), rm_confusion_matrix())
cv <- cv_standard(ds, cl, fps, rms, 3) 

# run a decoding analysis (this takes a few minutes) 
DECODING_RESULTS <- run_decoding(cv)
```


```{r line_plot, fig.height = 3, fig.width = 8, fig.align = "center"}
# plot the results for three different result types
plot(DECODING_RESULTS$rm_main_results, result_type = "all", plot_type = "line")
```



```{r TCD_plot, fig.height = 3.5, fig.width = 5, fig.align = "center"}
# create a temporal cross decoding plot
plot(DECODING_RESULTS$rm_main_results)
```




## Documentation

The documentation for this package is available at: https://emeyers.github.io/NDTr/


