---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```


# NeuroDecodeR: Neural Decoding in R


<!-- badges: start -->

[![Travis build
status](https://travis-ci.com/emeyers/NeuroDecodeR.svg?branch=master)](https://travis-ci.com/emeyers/NeuroDecodeR)

[![AppVeyor build
status](https://ci.appveyor.com/api/projects/status/github/emeyers/NeuroDecodeR?branch=master&svg=true)](https://ci.appveyor.com/project/emeyers/NeuroDecodeR)

[![Coveralls test
coverage](https://coveralls.io/repos/github/emeyers/NeuroDecodeR/badge.svg)](https://coveralls.io/r/emeyers/NeuroDecodeR?branch=master)

<!-- badges: end -->
  
<p>


## Overview

*Neural decoding* is a data analysis method that uses pattern classifiers to
predict experimental conditions based on neural activity. The Neural Decoding in
R (NDR) package makes it easy to do neural decoding analyses in R.


## Installation

You can install NeuroDecodeR package from github using:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("emeyers/NeuroDecodeR")
```


## Documentation

The documentation for this package is available at: https://emeyers.github.io/NeuroDecodeR/

To get started we recommend you read the [introductory tutorial](https://emeyers.github.io/NeuroDecodeR/articles/introduction_tutorial.html)


## Usage

The package is based on 5 abstract object types: 

1. `Datasources (DS)`: generate training and test sets.
2. `Feature preprocessors (FP)`: apply preprocessing to the training and test sets.
3. `Classifiers (CL)`: learn relationships on the training set and make
predictions on the test data.
4. `Result Metrics (RM)`: summarize the prediction accuracies.
5. `Cross-validators (CV)`: take the DS, FP and CL objects and run a
  cross-validation decoding procedure.

By combing different versions of these 5 object types together, it is possible
to run a range of different decoding analyses.

Below is a brief illustration of how to use the NDR to do a simple decoding
analysis. To learn how to use the NDR please see the [documentation
website](https://emeyers.github.io/NeuroDecodeR/) and the package vignettes. 


```{r basic_ndr_analysis}
library(NeuroDecodeR)

# file to data in "binned format"
basedir_file_name <- system.file(file.path("extdata", "ZD_150bins_50sampled.Rda"), package="NeuroDecodeR")

# create the DS, FP, CL, RM, and CV objects
ds <- ds_basic(basedir_file_name, 'stimulus_ID', 5, num_label_repeats_per_cv_split = 3)
fps <- list(fp_zscore())
cl <- cl_max_correlation()
rms <- list(rm_main_results(aggregate_normalized_rank = "diag"), rm_confusion_matrix())
cv <- cv_standard(ds, cl, fps, rms, 3) 

# run a decoding analysis (this takes a few minutes) 
DECODING_RESULTS <- run_decoding(cv)
```


```{r line_plot, fig.height = 3, fig.width = 8, fig.align = "center"}
# plot the results for three different result types
plot(DECODING_RESULTS$rm_main_results, result_type = "all", plot_type = "line")
```



```{r TCD_plot, fig.height = 3.5, fig.width = 5, fig.align = "center"}
# create a temporal cross decoding plot
plot(DECODING_RESULTS$rm_main_results)
```



## One can also run an analysis using the magrittr pipe (%>%) operator

```{r pipe_example, fig.height = 3, fig.width = 7, fig.align = "center"}
basedir_file_name <- system.file(file.path("extdata", "ZD_500bins_500sampled.Rda"), 
                                   package="NeuroDecodeR")
  
  DECODING_RESULTS <- basedir_file_name %>%
    ds_basic('stimulus_ID', 6, num_label_repeats_per_cv_split = 3) %>%
    cl_max_correlation() %>%
    fp_zscore() %>%
    rm_main_results() %>%
    rm_confusion_matrix() %>%
    cv_standard(num_resample_runs = 3) %>%
    run_decoding()
  
  plot(DECODING_RESULTS$rm_confusion_matrix)
```


