---
title: "Classifiers"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Classifiers}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---


Classifiers (CL) take a set of training data and training labels, and learn a model of the relationship between the training data and the labels from the different classes. Once this model has been learned (i.e., once the classifier has been trained), the classifier is then used to make predictions about what labels were present in a new set of ‘test data’. 


## Classifiers that come with the NDTr

There are three classifiers that come with the NDTr

* `cl_max_correation`
* `cl_poisson_naive_bayes`
* `cl_svm`


## Implementing a CL: required methods and data formats


Objects that are classifiers must implement the get_predictions() method. This method takes
two data frames called *training_set* and *test_set* have the following variables:

#### training_set

1. training_labels: The labels used to train the classifier
2. site_X's: a group of varaibles that has data from multiple sites

#### test_set

1. test_labels: The labels used to testthe classifier
2. site_X's: a group of varaibles that has data from multiple sites
3. time_bin: character strings listing which times different rows correspond to

The get_predictions() returns a data frame that has the following variables: 

1. test_time: a character vector indicating the times which the rows come from

2. actual_labels: the labels that were actually present on each trial

3. predicted_labels: the labels that the classifier predicted

4. decision_val.X (optional): a group of variables that has values that indicate how strongly
the classifier rates a test point as coming from a particular class


Like all NDTr objects, CL objects must also implement a get_properties() method
which returns a data frame with one row that lists all the properties that have
been set to allow for reproducible research.


## Example of internals of CL object using the cl_max_correlation


If you want to implement a new FP object yourself, below is an
example of how the FP object gets and returns data.


```{r, echo = FALSE, message=FALSE, warning=FALSE}
library(NDTr)
library(dplyr)
```


```{r}

# create a ds_basic to get the data
ds <- ds_basic("ZD_150bins_50sampled.Rda", 'stimulus_ID', 18)
cv_data <- get_data(ds)  
 

# an example of spliting the data into a training and test set, 
# this is done in the cross-validator
training_set <- dplyr::filter(cv_data, 
                              time_bin == "time.100_249", 
                              CV_1 == "train") %>%       # get data from the first CV split
  dplyr::select(starts_with("site"), train_labels)
        
test_set <- dplyr::filter(cv_data, CV_1 == "test") %>%   # get data from the first CV split
  dplyr::select(starts_with("site"), test_labels, time_bin) 



# use the cl object to make predictions 
cl <- cl_max_correlation()
predictions <- get_predictions(cl, training_set, test_set)

names(predictions)


# see how accurate the predictions are (chance is 1/7)
predictions_at_100ms <- dplyr::filter(predictions, test_time == "time.100_249")
mean(predictions_at_100ms$actual_labels == predictions_at_100ms$predicted_labels)


```













