---
title: "Datasources"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Datasources}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---


Datasources are used to generate training and tests splits of data. 



### Implementing a DS: required methods and data formats

All datasources must implement a get_data() method that returns a data frame that has the following variables in it: 

1. train_labels: The label levels that occur on each trial in the training data set

2. test_labels: The label levels that occur on each trial in the test data set

3. time_bin: The time in the experiment where the *test* data comes from

4. site_XXX: A collection of variables that each has data from one site  (e.g., neuron, EEG channel etc.) 

5. CV_XXX: A list for each CV split whether a given row is in that train or test set



### Example of internals of DS objects using the ds_basic object

Here is an example the data returned by the ds_basic datasource


```{r, echo = FALSE, warning=FALSE}
library(NDTr)
```


```{r}

ds <- ds_basic("ZD_150bins_50sampled.Rda", 'stimulus_ID', 18)

all_cv_data <- get_data(ds)  
 
names(all_cv_data)

```


<p>
