---
title: "Feature preprocessors"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Datasources}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---



## Feature Preprocessors

Feature preprocessors (FP) learn a set of parameters from the training data and
modify both the training and the test data based on these parameters, prior to
the data being sent to the classifier. The features preprocessor objects must
only use the training data to learn the preprocessing parameters in order to
prevent contamination between the training and test data which could bias the
results. 

### Implementing an FP: required methods and data formats

All feature preprocessors must implement preprocess_data(). This method takes
two data frames called *training_set* and *test_set* have the following
variables:

#### training_set

1. training_labels: The labels used to train the classifier
2. site_X's: a group of varaibles that has data from multiple sites

#### test_set

1. test_labels: The labels used to testthe classifier
2. site_X's: a group of varaibles that has data from multiple sites
3. time_bin: character strings listing which times different rows correspond to

The preprocess_data() returns a list with the two data frames *training_set* and
*test_set* but the data in these data frames has been preprocessed based on
parameters learned from the training_set


### Example of internals of FP objects using the fp_zscore

If you want to implement a new FP object yourself (advanced method), below is an
example of how the FP object gets and returns data.


```{r, echo = FALSE, message=FALSE, warning=FALSE}
library(NDTr)
library(dplyr)
```


```{r}

# create a ds_basic to get the data
ds <- ds_basic("ZD_150bins_50sampled.Rda", 'stimulus_ID', 18)
cv_data <- get_data(ds)  
 

# an example of spliting the data into a training and test set, 
# this is done in the cross-validator
training_set <- dplyr::filter(cv_data, 
                              time_bin == "time.100_249", 
                              CV_1 == "train") %>%       # get data from the first CV split
  dplyr::select(starts_with("site"), train_labels)
        
test_set <- dplyr::filter(cv_data, CV_1 == "test") %>%   # get data from the first CV split
  dplyr::select(starts_with("site"), test_labels, time_bin) 



# use the fp object to normalize the data 
fp <- fp_zscore()
processed_data <- preprocess_data(fp, training_set, test_set)

# prior to z-score normalizing the mean (e.g. for site 1) is not 0
mean(training_set$site_0001)

# after normalizing the data the mean is pretty much 0
mean(processed_data$training_set$site_0001)


```







